\section{Overview at the past attempts}

\subsection{LEP}

Although the hunt for the Higgs particle has been popularized recently with the completion of the LHC at CERN and the
\hbox{not-so-wild-black-hole-theories,} the hunt itself is a bit older.

They first started to search for it in LEP, CERN, which was an $e^-e^+$ collider. The dominant production process was $e^+e^-\to HZ$
~\cite{higgs:review}.
It was also possible to produce Higgs via other channels, but the cross sections were much lower. Those modes were secondary via
the $W^+W^-$ and $ZZ$ fusion~\cite{higgs:review}. The $Z$ in the $HZ$ production can be either virtual (LEP1 phase) or on mass shell
(LEP2 phase). Combined results were from center of mass energy $\sqrt{s} =$ 65 GeV to 209 GeV.

There were four four-jet decay modes that were particularly interesting:

\begin{itemize}
	\item[(a)] Final products are all quarks:
		\[ H \to b\bar{b}, \quad Z \to q\bar{q}, \] \vspace{-32pt}
	\item[(b)] Final products are quarks and tauons:
		\[ \left\{\begin{matrix}
		H \to \tau^+\tau^-, \quad Z \to q\bar{q}, \\
		H \to b\bar{b}, \quad Z \to \tau^+\tau^-,
		\end{matrix}\right.\] \vspace{-22pt}
	\item[(c)] Final products are quarks and neutrono pairs:
		\[ H \to b\bar{b}, \quad Z \to \nu\bar{\nu}, \] \vspace{-32pt}
	\item[(d)] Final products are light leptons and quarks:
		\[ H \to b\bar{b}, \quad Z \to \ell^+\ell^-, \] \vspace{-32pt} \label{higgs:decay}
\end{itemize}

where $q\bar{q}$ denotes quark--anti-quark pair, $\nu$ is a neutrino regardless of a generation, and $\ell^\pm \in \{e^\pm,\mu^\pm\}$.
LEP1 only used modes (c) and (d), while the LEP2 phase included all four of them.

LEP at the time had four collaborations working on trying to find Higgs: ALEPH, DELPHI, OPAL and L3. ALEPH found an excess of
$\sim 3 \sigma$, suggesting Higgs boson with mass $M_H \sim 115$ GeV, however others couldn't confirm such an excess, but then
again, they couldn't reject it at 95\% confidence level~\cite{higgs:review} (CL). Using this data they excluded existence of
SM Higgs boson below 114.4 GeV at 95\% CL, comparing data with theoretically calculated BR-s from Fig.~\ref{fig:SMHbrsmall}.

\vspace{12pt}

SM Higgs boson was also sought indirectly. Thus they obtained the upper limit
for the Higgs boson mass and at 95\% CL constrained it to be in the $114.4$ GeV $< M_H < 186$ GeV range\footnote{Accumulated data
of this indirect measurments over the last 20 years at LEP, SLC, Tevatron and elsewhere gave $M_H = 87^{+32}_{-26}$ GeV and
$M_H < 157$ GeV at 95\% CL.}, which coencides with theoretical predictions for metastable electroweak vacuum.

\subsection{Tevatron}

Two collaborations were working on the experiment: CDF and D\O. Main difference was that they Tevatron is $p\bar{p}$ collider and
because of this the production rates were much more diverse~\cite{tev4lhc} (Fig.~\ref{tev:prod}).

\begin{myfig}[11cm]{pics/tev.pdf}
	\vspace{-12pt}
	\caption{Tevatron SM Higgs boson nproduction rates, as calcluated by Tev4LHC Higgs working group~\cite{tev4lhc} on October
		2009. Higgs boson is marked with the `$h$' symbol.}
	\label{tev:prod}
\end{myfig}

Decay modes are already depicted on the fig.~\ref{fig:SMHbrsmall}.

Fig.~\ref{tev:prod} shows that the most promising channels are gluon fusion and production with vector bosons $Z,W^\pm$. QCD effects
give greater background and uncertainties are greater than at LEP. They tried to get rid of background using Monte Carlo
simulations for each specific channel and other calibrative experimental methods.

Each channel had to be checked differently:
\begin{itemize}
	\item $p\bar{p} \to W^\pm H$, where $H \to b\bar{b}$ and $W^\pm$ decays leptonically. Results are 4 to 5 times greater
		than the SM prediction for the $M_H$ of 115 GeV.
	\item $p\bar{p} \to ZH$, where, where the two decay like in the example (c) on page~\pageref{higgs:decay}. This required
		a very careful calculation of the missing transverse energy (neutrinos) with optimized triggers and analysis
		techniques.
	\item Same as above, but followed by the decay (d) on page~\pageref{higgs:decay}. This process had lesser $Z$ branching
		ratio, but the background was lower.
	\item $gg \to H$, $W^{\pm}H$, $ZH$ and then all decays with $H \to \tau^+\tau^-$. D\O{} also included a another one with
		$WH \to \tau\nu b\bar{b}$.
	\item All channels with $H \to \gamma\gamma$. This process is improbable within SM, but it sets strong constraints on
		possible SM extensions.
	\item Mixed search in CDF only with decay $WH + ZH \to jjb\bar{b}$, where $jj$ are any jet sprays coming from vector bosons.
		Because they used two modes at the same time they got greater signal, but then again the background was also
		greater.
\end{itemize}

They separated the hunt in two parts: low-mass Higgs with $M_H \leq 150$ GeV and high-mass with $M_H \geq 150$. Production rates
were 1.8 times greater for low-mass at  $M_H = 115$ GeV. Maximum was at $M_H \approx 135$ GeV, but it was below 2$\sigma$
~\cite[pg. 15]{higgs:review}. They managed to exclude the mass in the interval between 162 GeV and 166 GeV at 95\% CL.

\subsection{LHC}

LHC is a $pp$ collider. Again the theoretical calculations for production rates are represented by~\cite{tev4lhc}
(Fig~\ref{lhc:prod}).

\begin{myfig}{pics/lhc.pdf}
	\caption{Theoretical Higgs boson production cross sections for $pp \to HX$ at the courtesy of Tev4LHC Higgs working group,
		last updated on October 2009.}
	\label{lhc:prod}
\end{myfig}

Because of the monster energy range of the LHC they will try not only to search for Higgs bosons but also other more exotic
particles ($Z'$, \ldots). The hunt was divided into two energy ranges: below 130 GeV, where there is more background and several years
will be needed to accumulate sufficient statistics and above 130 GeV where they could confirm a discovery relatively quickly. Decay
modes that they are interested are mostly $H \to \gamma\gamma$, $H \to ZZ^* \to 4\ell$. Then vector boson fusion procecesses
resulting in $H \to \tau^+\tau^-, \gamma\gamma, W^+W^-$. Expected precision is $\sim 0.1\%$ in the 100 - 400 GeV mass range and
even for $M_H \sim 700$ GeV it would be $\sim 1\%$.

\vspace{12pt}

Upon the confirmed discovery it is planeed to check the Higgs boson for its basic quantum numbers (spin, CP-eigenvalue, coupling
constants for bosons and fermions).

At the LHC they also plan for a luminosity upgrade, called Super-LHC, which would also
allow for self-coupling measurments, because currently it's expected to be out of range.

Currently there has been an excess of 3.6~$\sigma$ for $M_H \sim 126$ GeV. With 95\% CL the $M_H$ has been restricted to be
$115.5$ GeV $< M_H < 131$ GeV.
