\section{Standard Model Higgs boson}

Contrary to the common (meaning layman) belief, Higgs bosons \emph{are} already a part of the Standard Model (SM).
In the sixties, there was a crisis in the world of particle physics: the theory all worked only for massless gauge bosons, and there
was also the hierarchy problem\footnote{This isn't really a problem, but merely an observation that the fundamental forces operate
at such a different scales. From here rose the question: is there any particular reason it is so?}. While such a mechanism cannot
give the direct answer to the latter\footnote{Though it is responsible for the electroweak symmetry breaking.}, it does provide the
masses for the gauge bosons.

Hypothesis is that our observable space is immersed in a field, which spontaneously breaks the symmetry of the Lagrangian density
$\mathcal{L}$. The original Lagrangian retains the symmetry (all gauge bosons and particles are massless), but the field is the one
that causes it.

There have been questions if the symmetry is only broken locally in space, or globally\footnote{If this field is inherently part of
the Minkowski space in which we currently appear to reside.}.

\subsection{The Higgs field}

Let us make a quick overview of some very basic theory behind the Higgs bosons just so we will get familiar with some of the
physical observables.

Just like in classical mechanics, Quantum Mechanics is also governed by the Lagrangian functions\cite{peskin}. We can write the
action as

\[
	S = \int \text{d}t\ L(q,\dot{q},t) = \int\text{d}t\int\text{d}^3x\ \mathcal{L}(q,\dot{q},t),
\]

where $S$ is action, $L$ our Lagrangean function and so $\mathcal{L}$ is the "density" of our Lagrangean $L$. In Minkowski space-time,
one can write

\[
	S = \int\text{d}^4x\ \mathcal{L}(q,\dot{q}),
\]

where $q$ is now a generalized set of coordinates.

We can obtain our set of our `equations of motion' by minimizing the action functional $S$. Just like in classical mechanics and in
the Theory of the Electromagnetic Field, we can substitute the spacial coordinates with fields themselves

\[
	\mathcal{L}(q,\dot{q}) \longrightarrow \mathcal{L}(\phi,\dot{\phi}),
\]

to get Euler-Lagrange equations, written in covariant form

\begin{equation}
	\frac{\partial\mathcal{L}}{\partial\phi} - \partial_\mu\left[\frac{\partial\mathcal{L}}{\partial(\partial_\mu\phi)}\right]
		= 0.
\end{equation}

Let us now consider the following $\mathcal{L}$ which describes the interaction between a scalar field $\phi$ and a Dirac fermion
$\psi$ \cite{wikipedia}:

\begin{equation}
	\mathcal{L} = \frac{1}{2}(\partial^\nu\phi)(\partial_\nu\phi) - \underbrace{(\mu^2\phi^2 + \lambda\phi^4)}_{V(\phi)} +
	\bar{\psi}(i\slashed{\partial} - m)\psi - \underbrace{g\ \bar{\psi}\phi\psi}_\text{Yukawa}.
\end{equation}

In this case $g$ is called the Yukawa coupling, and for $\mu < 0$ the symmetry of the $\mathcal{L}$ is spontaneously broken. This
means that $V(\phi)$ doesn't have its global minimum around 0, but around two opposing values $\phi_\pm$. If we now attack this
potential perurbatively, we must make the series expansion around one of those minumums. And the moment we decide which one to use,
we break the Lagrangean density symmetry: the field we observe then becomes, say, $\varphi = \phi - \phi_+$. In that case $\varphi$ is
called the Higgs field.

Because we deal with gravity here, the potential doesn't have the arbitrary vacuum state energy $v$. It's been theoretically set to
$v \approx 246$ GeV. It is depandant on the parameters $\mu$ and $\lambda$.
With more careful solving or simply using the Nambu-Goldstone theorem\footnote{Number of massless $N$-$G$ bosons equals to
$\text{dim} G - \text{dim} H = \# N\text{-}G$ bosons.}, we also get 3 massless Nambu-Goldstone bosons, that are absorbed to give
masses to the weak gauge bosons -- $Z, W^\pm$. Remaining massive component then becomes the Higgs boson, which is a "new player in
town".
\vspace{12pt}

While the Higgs field and the Higgs boson are a part of the SM, they still introduce new phyisics. Theorists are trying to set the
properties and find models using this fact so that they could fix the faults of the SM, such as absence of the dark matter candidate,
insufficient violation of CP symmetry (more about this in the further chapters).

